Keeping and discarding players

### 2020-05-27 inst/retain.Rmd
keep() and keep_flawed() implementation

---
title: "Illustrative use of retain()"
output: html_document
---

@@

## Function keep() vs function retain()

Functions `keep()` and `retain()` can be confusing, here is a tiny
example that should clarify things.

We will use a subset of
`skating_table`:

@@

```{r}
ordertable2supp(a,misslast=FALSE)
```

[note carefully how `misslast=FALSE` is needed here].  Now suppose we
are interested only in the first two competitors, `a` and `b`.  There
are two ways to discard competitor `c`, namely `keep()` and
`retain()`, which do slightly different things.  First, `keep()`:

```{r}
keep(ordertable2supp(a,misslast=FALSE),1:2)
```

Above, function `keep()` has effectively set `c=0` and then discarded
the `c^5` term.  This process, while natural, has the effect of
replacing `(a+c)^-4` with `a^-4`, which has no natural probabilistic
interpretation: we are ignoring `c`'s victories but nevertheless
interpreting joint strength term on the denominator [viz `a+c`] as a
*loss* for `a`.

```{r}
RA <- retain(a,1:2)
RA
```

@@

Above, object `or` is a perfectly sensible likelihood function.  What
happens if we `keep()` only `a`?

```{r}
keep(OR,1)
```

the above support function is meaningless (?) as, having set `b=0` we
then treat `a+b` as simply `a` which sort of makes sense until we
realise that `(a+b)^-6` reduces to `a^-6`.  It is as though `a` has
wins only 3 trials out of 8, against a competitor of known zero
strength whose wins we ignore.  Perhaps this has a natural
interpretation in probability, but if so I don't see it.

>>>>

---
title: "Keeping and discarding players"
output: html_document
---

@@

## Functions keep() and keep_flawed()

Functions `keep()` and `keep_flawed()` can be confusing, here is a
tiny example that should clarify things.  We will use a subset of
`skating_table`:

@@

```{r}
OT <- ordertable2supp(a,misslast=FALSE)
OT
```

[note carefully how `misslast=FALSE` is needed here].  Now suppose we
are interested only in the first two competitors, `a` and `b`.  There
are two ways to discard competitor `c`, namely `keep()` and
`keep_flawed()`, which do slightly different things.  First, `keep()`:

```{r}
keep_flawed(OT,1:2)
```

Above, function `keep_flawed()` has effectively taken `OT`, set `c=0`,
and then discarded the `c^5` term.  This process, while natural, has
the effect of replacing `(a+c)^-4` with `a^-4`, which has no natural
probabilistic interpretation: we are ignoring `c`'s victories but
nevertheless interpreting joint strength term on the denominator [viz
`a+c`] as a *loss* for `a`.  We will now use function `keep()`, which
operates on order tables:

```{r}
RA <- keep(a,1:2)
RA
```

@@

Above, object `or` is a perfectly sensible likelihood function.  What
happens if we keep only `a`?

```{r}
keep_flawed(OR,1)
```

the above support function is meaningless (?) as, having set `b=0` we
then treat `a+b` as simply `a` which sort of makes sense until we
realise that `(a+b)^-6` reduces to `a^-6`.  It is as though `a` has
wins only 3 trials out of 8, against a competitor of known zero
strength whose (impossible) wins we ignore.  Perhaps this has a
natural interpretation in probability, but if so I don't see it.
